### 2024-10-02 paper/proposal.qmd
set up initial files

---
title: "STA302 Final Project Proposal"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Author 2"
  - name: "Author 3"
thanks: "Code and data are available at: https://github.com/zzq20010617/sta302Paper"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
abstract: ""
number-sections: true
bibliography: proposal_references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)

```

# Introduction (350)

# Data description (300)
The R programming language (@talia) and ggplot2（@ggplot2）were used for processing the data.

# Ethics discussion (100-200)

# Preliminary results (300)

\newpage

# References

### 2024-10-02 scripts/00-download_data.R

@@ -1,0 +1,18 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from internet
+# Author: Ziqi Zhu
+# Date: 2 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: None
+
+
+#### Workspace setup ####
+library(readr)
+
+
+
+#### Save data ####
+write_csv(data, "data/raw_data/raw_data.csv") 
+
+         

### 2024-10-02 scripts/01-data_cleaning.R

@@ -1,0 +1,11 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded
+# Author: Ziqi Zhu
+# Date: 2 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+
+#### Clean data ####

### 2024-10-02 scripts/02-preliminary_data.R

@@ -1,0 +1,11 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziqi Zhu
+# Date: 2 October 2024
+# Contact: ziqi.zhu@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+
+#### preliminary data ####

### 2024-10-07 scripts/00-download_data.R
upload csv files and script for cleanning data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from internet
 # Author: Ziqi Zhu
-# Date: 2 October 2024
+# Date: 7 October 2024
 # Contact: ziqi.zhu@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: None
@@ -9,10 +9,18 @@
 
 #### Workspace setup ####
 library(readr)
-
+library(httr)
+library(jsonlite)
 
 
 #### Save data ####
+
+# download data from Kaggle
+system("kaggle datasets download -d nikdavis/steam-store-games -p data/raw_data")
+unzip("data/raw_data/steam-store-games.zip",exdir = "data/raw_data")
+
+# save data as raw_data.csv file
+data <- read.csv("data/raw_data/steam.csv")
 write_csv(data, "data/raw_data/raw_data.csv") 
 
          

### 2024-10-07 scripts/01-data_cleaning.R

@@ -1,11 +1,26 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded
 # Author: Ziqi Zhu
-# Date: 2 October 2024
+# Date: 7 October 2024
 # Contact: ziqi.zhu@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: None
 
 #### Workspace setup ####
+library(lubridate)
+library(dplyr)
 
 #### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+# Modify release date column and save as release month, select columns will be 
+# used for analysis save as analysus_data.csv
+cleaned_data <-
+  raw_data |> mutate(release_month = month(ymd(release_date),
+                                           label = TRUE, abbr = TRUE)) %>% 
+  select(release_month, english, developer, publisher, required_age, achievements,
+         positive_ratings, negative_ratings, average_playtime, median_playtime,
+         owners, price) %>%
+  filter(average_playtime != 0)
+
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-10-07 scripts/02-preliminary_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: 
-# Author: Ziqi Zhu
-# Date: 2 October 2024
+# Author: 
+# Date: 7 October 2024
 # Contact: ziqi.zhu@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: None

### 2024-10-09 paper/proposal.qmd
update data description section

@@
 library(ggplot2)
-
+library(readr)
+library(tibble)
+library(knitr)
+library(kableExtra)
+library(dplyr)
+data <- read.csv("../data/analysis_data/analysis_data.csv")
 ```
@@
 # Data description (300)
-The R programming language (@talia) and ggplot2（@ggplot2）were used for processing the data.
+We get the Steam games data(@davis2023steam) from Kaggle.
+The R programming language (@talia) and packages readr(@readr), httr(@httr), jsonlite(@jsonlite) were used to download the data, dplyr(@dplyr) and lubridate(@lubridate) were used to clean the data.
+Data was originally gathered from the Steam Store and SteamSpy APIs around May 2019.
+This table @tbl-cle shows the first 3 entries of cleaned data.
+We are going to take average_playtime as the response variable, it measures the mean time (in minutes) that players spend on a game, a summary of this variable is shown in @tbl-aveplaytime.
+This variable captures overall player engagement and serves as a good indicator of how immerse or entertaining a game is.
+It is suitable for a linear regression model because it is continuous and quantitative.
+The predictors selected are price, positive ratings, negative ratings, developer, and number of owners.
+All are numeric except for the number of owners, which is categorical since it represents estimated ranges.
+Price may influence playtime, as higher costs could lead to longer engagement.
+Positive ratings and negative ratings are reviews of players of that game and it can only be positive or negative.
+Positive ratings likely correlate with greater playtime, while negative ratings could indicate the opposite.
+Developer refers to the development company of the game, which could affect playtime, with established studios often producing longer, high-quality games.
+Lastly, the number of owners is an estimated number of owners, containing lower and upper bounds (like 20000-50000), it could signal popularity, where more owners may suggest higher average playtime.
+Summary of numerical predictors can be find in table @tbl-predictor.
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-cle
+#| tbl-cap: "Steam Games Data"
+
+head(data, 3) %>% 
+  kbl("latex") %>%
+  kable_styling(latex_options = c("striped", "scale_down")) 
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-aveplaytime
+#| fig-pos: "H"
+
+summary_table <- data %>%
+  summarise(
+    Mean = round(mean(average_playtime, na.rm = TRUE), 2),
+    Median = round(median(average_playtime, na.rm = TRUE), 2),
+    Std_Dev = round(sd(average_playtime, na.rm = TRUE), 2),
+    Min = round(min(average_playtime, na.rm = TRUE), 2),
+    Max = round(max(average_playtime, na.rm = TRUE), 2),
+  )
+kable(summary_table, caption = "Summary of Average Playtime")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-predictor
+
+predict_summary_table <- data %>%
+  summarise(
+    mean_price = round(mean(price, na.rm = TRUE), 2),
+    median_price = round(median(price, na.rm = TRUE), 2),
+    mean_positive_ratings = round(mean(positive_ratings, na.rm = TRUE), 2),
+    median_positive_ratings = round(median(positive_ratings, na.rm = TRUE), 2),
+    mean_negative_ratings = round(mean(negative_ratings, na.rm = TRUE), 2),
+    median_negative_ratings = round(median(negative_ratings, na.rm = TRUE), 2),
+  )
+kable(predict_summary_table, caption = "Summary of Predictors") %>%
+kable_styling(latex_options = c("striped", "scale_down"), font_size = 6) 
+```
 

### 2024-10-09 scripts/01-data_cleaning.R

@@ -18,7 +18,7 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 cleaned_data <-
   raw_data |> mutate(release_month = month(ymd(release_date),
                                            label = TRUE, abbr = TRUE)) %>% 
-  select(release_month, english, developer, publisher, required_age, achievements,
+  select(release_month, english, developer, publisher, achievements,
          positive_ratings, negative_ratings, average_playtime, median_playtime,
          owners, price) %>%
   filter(average_playtime != 0)

### 2024-10-11 paper/proposal.qmd
add introduction

@@
 # Introduction (350)
+The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam.
+Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers.
+Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts.
+So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle.
+Specifically, we will investigate how variables such as price, user ratings, and other game characteristics contribute to player engagement, as measured by playtime.
+From some previous papers, we learned that both positive and negative reviews have a significant correlation with playtime, but the effect is more pronounced for positive reviews(@paper1).
+Another paper found out that game pricing significantly affects player engagement, with lower-priced games generally seeing higher playtime.(@paper2).
+The third paper did an empirical study that also examined the influence of game reviews on playtime, focusing on how the number and sentiment of reviews can affect player engagement.
+It concludes that games with more positive reviews see increased playtime, especially if those reviews highlight the game’s quality(@paper3).
+We found this problem suitable for linear regression because it allows us to model the relationship between average playtime (the dependent variable) and various predictors (price, ratings, developer, etc.).
+This approach helps quantify how each factor influences playtime, enabling a clear understanding of the impact of different game attributes.
+By fitting a linear regression model, we can determine which factors have statistically significant effects on playtime, providing actionable insights for game developers and marketers.
 
@@
 #| tbl-cap: "Steam Games Data"
+#| tbl-float: false
 

### 2024-10-11 paper/proposal.qmd
Refactored ethics discussion

--- 2024-10-09 paper/proposal.qmd
@@
 # Ethics discussion (100-200)
-
+The Steam Games dataset used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform.
+The metadata is comprehensively filled out, including details such as game titles, release dates, genres, and user ratings, which enhances the dataset's usability and reliability.
+The source of the data is clearly described, originating from the Steam store pages and Steam API, ensuring transparency and traceability.
+Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, which implies a level of vetting and popularity within the data science community.
+This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility for academic research.
 # Preliminary results (300)

### 2024-10-11 paper/proposal.qmd
Refactored ethics discussion

@@
 # Ethics discussion (100-200)
-The Steam Games dataset used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform.
+The Steam Games dataset(@davis2023steam) used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform.
 The metadata is comprehensively filled out, including details such as game titles, release dates, genres, and user ratings, which enhances the dataset's usability and reliability.
@@
 Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, which implies a level of vetting and popularity within the data science community.
-This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility for academic research.
+This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility.
+There are no ethical concerns to report.
 # Preliminary results (300)

### 2024-10-11 paper/proposal.qmd
Update proposal.qmd

@@
 
+# Fit multiple linear regression model
+model <- lm(average_playtime ~ price + positive_ratings + negative_ratings + developer + owners)
+
+# Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+
+y_value <- resid(model)
+x_value <- fitted(model)
+
+plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
+
+# Create Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+
+par(mfrow=c(1,3))
+
+plot(x = data$price, y = y_value, main = "Residuals vs. Price", xlab = "Price", ylab = "Residuals")
+
+plot(x = data$positive_ratings, y = y_value, main = "Residuals vs. Positive Ratings", xlab = "Positive Ratings", ylab = "Residuals")
+
+plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")
+
+# Create BoxPlot to check assumptions for Categorical Predictors
+
+par(mfrow=c(1,2))
+
+boxplot(y_value ~ data$developer , main="Residual vs Developer",
+xlab="Developer", ylab="Residuals")
+
+boxplot(y_value ~ data$owners , main="Residual vs Owners",
+xlab="Owners", ylab="Residuals")
+
+# Create Normal QQ Plot to check Normality assumption
+
+qqnorm(y_value)
+qqline(y_value)
+
 \newpage

### 2024-10-11 paper/proposal.qmd
Update proposal.qmd

@@
 
-# Fit multiple linear regression model
+## Fit multiple linear regression model
 model <- lm(average_playtime ~ price + positive_ratings + negative_ratings + developer + owners)
 
-# Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
 
@@
 
-# Create Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+## Create Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
 
@@
 
-# Create BoxPlot to check assumptions for Categorical Predictors
+## Create BoxPlot to check assumptions for Categorical Predictors
 
@@
 
-# Create Normal QQ Plot to check Normality assumption
+## Create Normal QQ Plot to check Normality assumption
 

### 2024-10-11 paper/proposal.qmd
generate model and graph

--- 2024-10-11 paper/proposal.qmd
@@
 There are no ethical concerns to report.
+
 # Preliminary results (300)
+
+## Fit multiple linear regression model
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+model <- lm(average_playtime ~ price + positive_ratings + negative_ratings + developer + owners, data=data)
+```
+
+## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+y_value <- resid(model)
+x_value <- fitted(model)
+
+plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
+```
+
+## Create Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+par(mfrow=c(1,3))
+
+plot(x = data$price, y = y_value, main = "Residuals vs. Price", xlab = "Price", ylab = "Residuals")
+
+plot(x = data$positive_ratings, y = y_value, main = "Residuals vs. Positive Ratings", xlab = "Positive Ratings", ylab = "Residuals")
+
+plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")
+```
+## Create BoxPlot to check assumptions for Categorical Predictors
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+par(mfrow=c(1,2))
+
+boxplot(y_value ~ data$developer , main="Residual vs Developer",
+xlab="Developer", ylab="Residuals")
+
+boxplot(y_value ~ data$owners , main="Residual vs Owners",
+xlab="Owners", ylab="Residuals")
+```
+## Create Normal QQ Plot to check Normality assumption
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+qqnorm(y_value)
+qqline(y_value)
+```
 

### 2024-10-11 paper/proposal.qmd
Refactored preliminary analysis to include pairwise and response vs fitted plots

@@
 library(ggplot2)
+library(GGally)
 library(readr)
@@
 ```
+## Create Response vs Fitted Scatterplot to check patterns in residuals before further analysis.
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+y_hat <- fitted(model)
+plot(x = y_hat, y = data$average_playtime, main = "Response vs. Fitted", xlab = "Fitted", ylab = "Average Playtime")
+abline(a=0, b=1, lty=2)
+```
+## Create Pairwise Scatterplots for Predictors to look for multicollinearity
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+ggpairs(data[, c(6, 7, 10, 11)])
 
+```
 ## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance

### 2024-10-11 paper/proposal.qmd
Update proposal.qmd

@@
 ```
-
+The four linear regression assumptions to be looked into are linearity, uncorrelated errors, constant error variance, and normality.
+Since this is a multiple linear regression, we must also check the conditional mean response and predictor conditions.
+To check the conditional mean response condition, we look at the Response vs.
+Fitted scatterplot.
+We see an that the data points are clustered fairly tightly in the bottom left corner of the scatterplot, which satisfies the condition.
+To check the conditional mean predictors assumption, we must look at the pairwise scatterplots of predictors, in which we see a lack of curves which satisfies this condition as well.
+When looing at the Residual vs.
+Fitted scatterplot, we can see a fanning pattern in the shape of a left pointing arrow, which leads to a possible violation of the constant variance assumption.
+Looking at the Residuals vs.
+Qualitative Predictor scatterplots suggests a possible violation of both linearity and uncorrelated errors due to the cluster of many points in all three as well as the curves that exist in the positive and negative ratings.
+Neither of the boxplots appear to suggest any kind of linear assumption violation regarding the categorical predictors.
+Looking at the Normal QQ Plot, there is some deviation from the diagonal at the extremes that appears to be too significant to ignore.
+This leads to a normality violation.
 \newpage

### 2024-10-11 paper/proposal.qmd
Update proposal.qmd

@@
 This leads to a normality violation.
+There being so many possible assumption violations makes sense due to the nature of the data.
+For example, the violation of the uncorrelated errors assumption is valid because the predictors would depend on one another if even slightly (i.e.
+price and developer).
+The results of this preliminary model are fairly similar to the results in the literature as similar errors and violated assumptions were noted.
+Transformations in order to mitigate these violations will be vital to maintain the stability of the normal model.
+
 \newpage

### 2024-10-11 paper/proposal.Rmd
add rmd file, add contributions

--- 2024-10-11 paper/proposal.qmd
@@
   - name: "Ziqi Zhu"
-  - name: "Author 2"
-  - name: "Author 3"
+  - name: "Raghav Sinha"
+  - name: "Christophe McWatt"
 thanks: "Code and data are available at: https://github.com/zzq20010617/sta302Paper"
@@
 
-# Introduction (350)
+# Contributions
+We found the dataset together, and seperate the works as follow:
+Ziqi: Introduction and data description
+Raghav: Ethics discussion and some plot in preliminary result
+Christophe: Most of Preliminary results section
+
+# Introduction
 The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam.
@@
 
-# Data description (300)
+# Data description
 We get the Steam games data(@davis2023steam) from Kaggle.
@@
 Positive ratings likely correlate with greater playtime, while negative ratings could indicate the opposite.
-Developer refers to the development company of the game, which could affect playtime, with established studios often producing longer, high-quality games.
+Developer refers to the development company of the game, which could affect playtime, with established studios often producing longer, high-quality games, but since they are too many developers to fit we are planning to create bins or indicator variables in later phase.
 Lastly, the number of owners is an estimated number of owners, containing lower and upper bounds (like 20000-50000), it could signal popularity, where more owners may suggest higher average playtime.
@@
 
-# Ethics discussion (100-200)
+# Ethics discussion
 The Steam Games dataset(@davis2023steam) used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform.
@@
 
-# Preliminary results (300)
+# Preliminary results
 

### 2024-10-11 paper/proposal.Rmd
add rmd file, add contributions

@@
 This leads to a normality violation.
+There being so many possible assumption violations makes sense due to the nature of the data.
+For example, the violation of the uncorrelated errors assumption is valid because the predictors would depend on one another if even slightly (i.e.
+price and developer).
+The results of this preliminary model are fairly similar to the results in the literature as similar errors and violated assumptions were noted.
+Transformations in order to mitigate these violations will be vital to maintain the stability of the normal model.
 \newpage

### 2024-11-27 paper/proposal.Rmd
Amended ethics discussion according to feedback from P1

@@
 The source of the data is clearly described, originating from the Steam store pages and Steam API, ensuring transparency and traceability.
-Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, which implies a level of vetting and popularity within the data science community.
-This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility.
+Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, and has been vetted and used in existing literature (discussed in the introduction).
+This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility through the framework outlined in the ethics module.
 There are no ethical concerns to report.

### 2024-11-27 paper/proposal.tex
Amended ethics discussion according to feedback from P1

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={STA302 Final Project Proposal},
  pdfauthor={Ziqi Zhu; Author 2; Author 3},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{STA302 Final Project Proposal\thanks{Code and data are available
at: https://github.com/zzq20010617/sta302Paper}}
\author{Ziqi Zhu \and Author 2 \and Author 3}
\date{October 23, 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, sharp corners, frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{introduction-350}{%
\section{Introduction (350)}\label{introduction-350}}

The gaming industry has experienced exponential growth in recent years,
with millions of players engaging in various games on platforms like
Steam. Understanding what influences average playtime in games is
crucial for game developers, publishers, and marketers. Average playtime
is a key indicator of player engagement and game success, and
investigating factors that impact it can provide valuable insights into
game design, pricing strategies, and marketing efforts. So our research
aims to explore the factors that influence the average playtime of games
available on Steam using a dataset we get from Kaggle. Specifically, we
will investigate how variables such as price, user ratings, and other
game characteristics contribute to player engagement, as measured by
playtime. From some previous papers, we learned that both positive and
negative reviews have a significant correlation with playtime, but the
effect is more pronounced for positive reviews(Brodschneider and Pirker
(2023)). Another paper found out that game pricing significantly affects
player engagement, with lower-priced games generally seeing higher
playtime.(Luisa et al. (2021)). The third paper did an empirical study
that also examined the influence of game reviews on playtime, focusing
on how the number and sentiment of reviews can affect player engagement.
It concludes that games with more positive reviews see increased
playtime, especially if those reviews highlight the game's quality(Lin
et al. (2019)). We found this problem suitable for linear regression
because it allows us to model the relationship between average playtime
(the dependent variable) and various predictors (price, ratings,
developer, etc.). This approach helps quantify how each factor
influences playtime, enabling a clear understanding of the impact of
different game attributes. By fitting a linear regression model, we can
determine which factors have statistically significant effects on
playtime, providing actionable insights for game developers and
marketers.

\hypertarget{data-description-300}{%
\section{Data description (300)}\label{data-description-300}}

We get the Steam games data(Davis (2023)) from Kaggle. The R programming
language (R Core Team (2023)) and packages readr(Wickham, Hester, and
Bryan (2024)), httr(Wickham (2023)), jsonlite(Ooms (2014)) were used to
download the data, dplyr(Wickham et al. (2023)) and lubridate(Grolemund
and Wickham (2011)) were used to clean the data. Data was originally
gathered from the Steam Store and SteamSpy APIs around May 2019. This
table Table~\ref{tbl-cle} shows the first 3 entries of cleaned data. We
are going to take average\_playtime as the response variable, it
measures the mean time (in minutes) that players spend on a game, a
summary of this variable is shown in Table~\ref{tbl-aveplaytime}. This
variable captures overall player engagement and serves as a good
indicator of how immerse or entertaining a game is. It is suitable for a
linear regression model because it is continuous and quantitative. The
predictors selected are price, positive ratings, negative ratings,
developer, and number of owners. All are numeric except for the number
of owners, which is categorical since it represents estimated ranges.
Price may influence playtime, as higher costs could lead to longer
engagement. Positive ratings and negative ratings are reviews of players
of that game and it can only be positive or negative. Positive ratings
likely correlate with greater playtime, while negative ratings could
indicate the opposite. Developer refers to the development company of
the game, which could affect playtime, with established studios often
producing longer, high-quality games. Lastly, the number of owners is an
estimated number of owners, containing lower and upper bounds (like
20000-50000), it could signal popularity, where more owners may suggest
higher average playtime. Summary of numerical predictors can be find in
table Table~\ref{tbl-predictor}.

\hypertarget{tbl-cle}{}
\begin{table}
\caption{\label{tbl-cle}Steam Games Data }\tabularnewline

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{l|r|l|l|r|r|r|r|r|l|r}
\hline
release\_month & english & developer & publisher & achievements & positive\_ratings & negative\_ratings & average\_playtime & median\_playtime & owners & price\\
\hline
\cellcolor{gray!10}{Nov} & \cellcolor{gray!10}{1} & \cellcolor{gray!10}{Valve} & \cellcolor{gray!10}{Valve} & \cellcolor{gray!10}{0} & \cellcolor{gray!10}{124534} & \cellcolor{gray!10}{3339} & \cellcolor{gray!10}{17612} & \cellcolor{gray!10}{317} & \cellcolor{gray!10}{10000000-20000000} & \cellcolor{gray!10}{7.19}\\
\hline
Apr & 1 & Valve & Valve & 0 & 3318 & 633 & 277 & 62 & 5000000-10000000 & 3.99\\
\hline
\cellcolor{gray!10}{May} & \cellcolor{gray!10}{1} & \cellcolor{gray!10}{Valve} & \cellcolor{gray!10}{Valve} & \cellcolor{gray!10}{0} & \cellcolor{gray!10}{3416} & \cellcolor{gray!10}{398} & \cellcolor{gray!10}{187} & \cellcolor{gray!10}{34} & \cellcolor{gray!10}{5000000-10000000} & \cellcolor{gray!10}{3.99}\\
\hline
\end{tabular}}
\end{table}

\begin{table}

\caption{\label{tbl-aveplaytime}\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-aveplaytime-1}Summary of Average Playtime}

{\centering

\begin{tabular}[t]{rrrrr}
\toprule
Mean & Median & Std\_Dev & Min & Max\\
\midrule
657.37 & 222 & 3783.67 & 1 & 190625\\
\bottomrule
\end{tabular}

}

\end{minipage}%

\end{table}

\begin{table}

\caption{\label{tbl-predictor}\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-predictor-1}}

{\centering

\begingroup\fontsize{6}{8}\selectfont

}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-predictor-2}Summary of Predictors}

{\centering

\begin{longtable}[t]{rrrrrr}
\\
\toprule
mean\_price & median\_price & mean\_positive\_ratings & median\_positive\_ratings & mean\_negative\_ratings & median\_negative\_ratings\\
\midrule
\cellcolor{gray!10}{7.47} & \cellcolor{gray!10}{4.99} & \cellcolor{gray!10}{4181.49} & \cellcolor{gray!10}{408} & \cellcolor{gray!10}{858.13} & \cellcolor{gray!10}{113}\\
\bottomrule
\end{longtable}
\endgroup{}

}

\end{minipage}%

\end{table}

\hypertarget{ethics-discussion-100-200}{%
\section{Ethics discussion (100-200)}\label{ethics-discussion-100-200}}

The Steam Games dataset(Davis (2023)) used in this research is a
collected dataset, not simulated, comprising information on over 97,000
games published on the Steam platform. The metadata is comprehensively
filled out, including details such as game titles, release dates,
genres, and user ratings, which enhances the dataset's usability and
reliability. The source of the data is clearly described, originating
from the Steam store pages and Steam API, ensuring transparency and
traceability. Additionally, the dataset is hosted on Kaggle, a reputable
platform for data science and machine learning, which implies a level of
vetting and popularity within the data science community. This
widespread use and accessibility suggest that the dataset is
well-regarded and trusted by third parties, further validating its
credibility. There are no ethical concerns to report.

\hypertarget{preliminary-results-300}{%
\section{Preliminary results (300)}\label{preliminary-results-300}}

\hypertarget{fit-multiple-linear-regression-model}{%
\subsection{Fit multiple linear regression
model}\label{fit-multiple-linear-regression-model}}

\hypertarget{create-response-vs-fitted-scatterplot-to-check-patterns-in-residuals-before-further-analysis.}{%
\subsection{Create Response vs Fitted Scatterplot to check patterns in
residuals before further
analysis.}\label{create-response-vs-fitted-scatterplot-to-check-patterns-in-residuals-before-further-analysis.}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-6-1.pdf}

\hypertarget{create-pairwise-scatterplots-for-predictors-to-look-for-multicollinearity}{%
\subsection{Create Pairwise Scatterplots for Predictors to look for
multicollinearity}\label{create-pairwise-scatterplots-for-predictors-to-look-for-multicollinearity}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-7-1.pdf}

\hypertarget{create-residuals-vs-fitted-scatterplot-to-check-linearity-uncorrelated-errors-and-constant-variance}{%
\subsection{Create Residuals vs Fitted Scatterplot to check Linearity,
Uncorrelated Errors, and Constant
Variance}\label{create-residuals-vs-fitted-scatterplot-to-check-linearity-uncorrelated-errors-and-constant-variance}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-8-1.pdf}

\hypertarget{create-residuals-vs-each-quantitative-predictor-scatterplot-to-check-linearity-uncorrelated-errors-and-constant-variance}{%
\subsection{Create Residuals vs each Quantitative Predictor scatterplot
to check Linearity, Uncorrelated Errors, and Constant
Variance}\label{create-residuals-vs-each-quantitative-predictor-scatterplot-to-check-linearity-uncorrelated-errors-and-constant-variance}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-9-1.pdf}

\hypertarget{create-boxplot-to-check-assumptions-for-categorical-predictors}{%
\subsection{Create BoxPlot to check assumptions for Categorical
Predictors}\label{create-boxplot-to-check-assumptions-for-categorical-predictors}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-10-1.pdf}

\hypertarget{create-normal-qq-plot-to-check-normality-assumption}{%
\subsection{Create Normal QQ Plot to check Normality
assumption}\label{create-normal-qq-plot-to-check-normality-assumption}}

\includegraphics{proposal_files/figure-pdf/unnamed-chunk-11-1.pdf}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-paper1}{}}%
Brodschneider, Vinzenz, and Johanna Pirker. 2023. {``On the Influence of
Reviews on Play Activity on Steam - a Statistical Approach.''}
\emph{ResearchGate}.
\url{https://www.researchgate.net/publication/376227235_On_the_Influence_of_Reviews_on_Play_Activity_on_Steam_-_A_Statistical_Approach}.

\leavevmode\vadjust pre{\hypertarget{ref-davis2023steam}{}}%
Davis, Nik. 2023. {``Steam Store Games (Clean Dataset).''}
\url{https://www.kaggle.com/datasets/nikdavis/steam-store-games/data}.

\leavevmode\vadjust pre{\hypertarget{ref-lubridate}{}}%
Grolemund, Garrett, and Hadley Wickham. 2011. {``Dates and Times Made
Easy with {lubridate}.''} \emph{Journal of Statistical Software} 40 (3):
1--25. \url{https://www.jstatsoft.org/v40/i03/}.

\leavevmode\vadjust pre{\hypertarget{ref-paper3}{}}%
Lin, Dayi, Cor-Paul Bezemer, Ahmed E. Hassan, and Ying Zou. 2019. {``An
Empirical Study of Game Reviews on the Steam Platform.''} In \emph{Empir
Software Eng}, 24.
\url{https://doi-org.myaccess.library.utoronto.ca/10.1007/s10664-018-9627-4}.

\leavevmode\vadjust pre{\hypertarget{ref-paper2}{}}%
Luisa, Andraž De, Jan Hartman, David Nabergoj, and Samo Pahor. 2021.
{``Predicting the Popularity of Games on Steam.''} \emph{ResearchGate}.
\url{https://www.researchgate.net/publication/355110719_Predicting_the_Popularity_of_Games_on_Steam}.

\leavevmode\vadjust pre{\hypertarget{ref-jsonlite}{}}%
Ooms, Jeroen. 2014. {``The Jsonlite Package: A Practical and Consistent
Mapping Between JSON Data and r Objects.''} \emph{arXiv:1403.2805
{[}Stat.CO{]}}. \url{https://arxiv.org/abs/1403.2805}.

\leavevmode\vadjust pre{\hypertarget{ref-talia}{}}%
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-httr}{}}%
Wickham, Hadley. 2023. \emph{Httr: Tools for Working with URLs and
HTTP}. \url{https://CRAN.R-project.org/package=httr}.

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
Vaughan. 2023. \emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\vadjust pre{\hypertarget{ref-readr}{}}%
Wickham, Hadley, Jim Hester, and Jennifer Bryan. 2024. \emph{Readr: Read
Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\end{CSLReferences}

\end{document}

### 2024-11-27 paper/proposal.Rmd
Removed 'owners' from pairplot and made owners and owners_combined factored columns

@@
 #| message: false
-ggpairs(data[, c(6, 7, 10, 11)])
+
+ggpairs(data[, c(6, 7, 11)]) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+        axis.text.y = element_text(size = 8))
 

### 2024-11-27 scripts/01-data_cleaning.R
Changed owners and owners_combined to factored columns

@@ -9,18 +9,42 @@
 #### Workspace setup ####
 library(lubridate)
 library(dplyr)
+library(readr)
 
 #### Clean data ####
+# Read the raw data
 raw_data <- read_csv("data/raw_data/raw_data.csv")
 
-# Modify release date column and save as release month, select columns will be 
-# used for analysis save as analysus_data.csv
-cleaned_data <-
-  raw_data |> mutate(release_month = month(ymd(release_date),
-                                           label = TRUE, abbr = TRUE)) %>% 
+# Clean and transform the data
+cleaned_data <- raw_data %>%
+  mutate(
+    release_month = month(ymd(release_date), label = TRUE, abbr = TRUE),
+    owners_combined = recode(owners,
+                             "5000000-10000000" = "5000000-20000000",
+                             "10000000-20000000" = "5000000-20000000",
+                             "20000000-50000000" = "20000000-200000000",
+                             "100000000-200000000" = "20000000-200000000",
+                             "50000000-100000000" = "20000000-200000000",
+                             .default = as.character(owners))
+  ) %>%
   select(release_month, english, developer, publisher, achievements,
          positive_ratings, negative_ratings, average_playtime, median_playtime,
-         owners, price) %>%
+         owners, price, owners_combined) %>%
   filter(average_playtime != 0)
 
+# Define the order of levels for the owners and owners_combined columns
+levels_order <- c("0-20000", "20000-50000", "50000-100000", "100000-200000", 
+                  "200000-500000", "500000-1000000", "1000000-2000000", 
+                  "2000000-5000000", "5000000-10000000", "10000000-20000000", 
+                  "20000000-50000000", "50000000-100000000", "100000000-200000000")
+
+levels_order_combined <- c("0-20000", "20000-50000", "50000-100000", "100000-200000", 
+                           "200000-500000", "500000-1000000", "1000000-2000000", 
+                           "2000000-5000000", "5000000-20000000", "20000000-200000000")
+
+# Apply the factor levels
+cleaned_data$owners <- factor(cleaned_data$owners, levels = levels_order)
+cleaned_data$owners_combined <- factor(cleaned_data$owners_combined, levels = levels_order_combined)
+
+# Save the cleaned data
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-11-30 paper/proposal.Rmd
Implemented concepts up until module 8

@@
 
-model <- lm(average_playtime ~ price + positive_ratings + negative_ratings + developer + owners, data=data)
+model <- lm(average_playtime ~ price + positive_ratings + negative_ratings + achievements + owners, data=data)
 ```
@@
 
-ggpairs(data[, c(6, 7, 11)]) +
+ggpairs(data[, c(5, 6, 7, 11)]) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
@@
 #| message: false
-par(mfrow=c(1,3))
+par(mfrow=c(1,4))
 
@@
 plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")
+
+plot(x = data$achievements, y = y_value, main = "Residuals vs. Achievements", xlab = "Achievements", ylab = "Residuals")
 ```
@@
 ```
+## Box-Cox Transformation
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+library(MASS)
+library(caret)
+
+# Define the Box-Cox transformation function
+boxcox_transform <- function(x) {
+  bc <- BoxCoxTrans(x)
+  predict(bc, x)
+}
+
+# Apply the transformation to each predictor and the response
+bc_average_playtime <- boxcox_transform(data$average_playtime)
+bc_price <- boxcox_transform(data$price)
+bc_positive_ratings <- boxcox_transform(data$positive_ratings)
+bc_negative_ratings <- boxcox_transform(data$negative_ratings)
+bc_achievements <- boxcox_transform(data$achievements)
+```
+## Fitting transformed model
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+model_transformed <- lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements+ bc_negative_ratings + owners, data=data)
+```
+## Checking for Multicollinearity
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+# install.packages("car")
+library(car)
+vif(model_transformed)
+```
+## Fitting Reduced Model
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+reduced <- lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)
+summary(reduced)
+```
+## Generatingn Residual Plots
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+par(mfrow = c(2, 2))
+plot(reduced)
+```
+## AIC Model Selection
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+step_model <- stepAIC(reduced, direction = "both")
+summary(step_model)
+```
+## ANOVA Comparison between full model and reduced model
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+anova(step_model, reduced)
+# Conclude that step_model (without bc_achievements is better suited and no significant improvement to model from bc_achievements)
+```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+summary(step_model)
+```
 The four linear regression assumptions to be looked into are linearity, uncorrelated errors, constant error variance, and normality.
@@
 To check the conditional mean predictors assumption, we must look at the pairwise scatterplots of predictors, in which we see a lack of curves which satisfies this condition as well.
-When looing at the Residual vs.
+When looking at the Residual vs.
 Fitted scatterplot, we can see a fanning pattern in the shape of a left pointing arrow, which leads to a possible violation of the constant variance assumption.

### 2024-12-01 paper/paper.Rmd
update paper, save models in rds files

---
title: "Exploring Factors Influencing Average Playtime of Video Games: A Regression Analysis"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Raghav Sinha"
  - name: "Christophe McWatt"
thanks: "Code and data are available at: [https://github.com/zzq20010617/sta302Paper](https://github.com/zzq20010617/sta302Paper)."
date: today
date-format: long
abstract: ""
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(GGally)
library(readr)
library(tibble)
library(knitr)
library(kableExtra)
library(dplyr)
library(MASS)
library(caret)
library(car)
library(broom)
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load analysis data
data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Contributions
We found the dataset together, and seperate the works as follow:
Ziqi: Introduction and data description
Raghav: Ethics discussion and some plot in preliminary result
Christophe: Most of Preliminary results section

# Introduction
The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam. Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers. Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts. So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle. Specifically, we will investigate how variables such as price, user ratings, and other game characteristics contribute to player engagement, as measured by playtime.
From some previous papers, we learned that both positive and negative reviews have a significant correlation with playtime, but the effect is more pronounced for positive reviews [@paper1]. Another paper found out that game pricing significantly affects player engagement, with lower-priced games generally seeing higher playtime [@paper2]. The third paper did an empirical study that also examined the influence of game reviews on playtime, focusing on how the number and sentiment of reviews can affect player engagement. It concludes that games with more positive reviews see increased playtime, especially if those reviews highlight the game’s quality [@paper3].
We found this problem suitable for linear regression because it allows us to model the relationship between average playtime (the dependent variable) and various predictors (price, ratings, developer, etc.). This approach helps quantify how each factor influences playtime, enabling a clear understanding of the impact of different game attributes. By fitting a linear regression model, we can determine which factors have statistically significant effects on playtime, providing actionable insights for game developers and marketers.

# Methods

To develop and refine the final model for predicting average playtime, a structured approach was employed that combined diagnostic assessments, transformations, and model selection techniques. The analysis began with data preparation and selection of relevant predictors based on domain knowledge and preliminary exploration. A multiple linear regression model was initially fitted using predictors including price, positive ratings, negative ratings, number of achievements as numeric variables, and ownership levels as categorical variable.

Preliminary model diagnostics, since we are using MLR model, we first need to check conditional mean condition with both response and predictors, including response vs. fitted plots and pairwise plots between predictors. Then residuals vs. fitted plots were used to assess the assumptions of linearity, constant-variance, and uncorrelated errors. While Normal Quantile-Quantile (QQ) plot is used to check for normality. To address violations of these assumptions, Box-Cox transformations were applied to the response variable and select predictors to make the most suited transform function based on their $\lambda$. Then pairwise scatterplots and the Variance Inflation Factor (VIF) were used to check for multicollinearity, and highly correlated variables were removed or adjusted to improve the model's stability.

Refinement of the model was carried out through stepwise selection using AIC (Akaike Information Criterion) to identify the best combination of predictors while balancing model complexity and fit. ANOVA tests were used to compare nested models and evaluate whether reductions in predictors led to significant changes in explanatory power. The final model was validated using diagnostic plots to confirm adherence to linear regression assumptions, ensuring that it was suitable for interpretation and prediction.

The methodology was implemented in programming language **R** [@citeR]. Packages **readr** [@readr], **httr** [@httr], **jsonlite** [@jsonlite] were used to download the data, **dplyr** [@dplyr] and **lubridate** [@lubridate] were used to clean the data. The diagnostic process is done with help of **ggplot2** [@ggplot2] for residual graphs, **GGally** [@GGally] and **car** [@car] for multicollinearity checks, **MASS** [@MASS] for stepwise selection, **caret** [@caret] is used to perform box-cox transform on variables. This systematic approach ensured that each decision was guided by statistical evidence and diagnostic tools, allowing for the development of a robust final model. A simplified version of the methodology flowchart is presented in @fig-flowchart.

```{r}
#| warning: false
#| message: false
#| echo: false
#| out.width: 100%
#| label: fig-flowchart
#| fig-cap: "Methodology flowchart"

include_graphics(here::here("other/flowchart.png"))
```

<!-- ## Fit multiple linear regression model -->
```{r}
#| echo: false
#| warning: false
#| message: false

# Load first model fit with lm(average_playtime ~ price + positive_ratings + 
# negative_ratings + achievements + owners, data=analysis_data)
model_1 <- readRDS(file = here::here("models/model_1.rds"))
```
# Results

At the starting point, our baseline model uses `average_playtime` as the response variable, with `price`, `positive_ratings`, `negative_ratings`, `achievements`, and `owners` as predictors. The four key linear regression assumptions considered are linearity, uncorrelated errors, constant error variance, and normality. Additionally, since this is a multiple linear regression model, we also examine the conditional mean response and predictor assumptions.
<!-- ## Create Response vs Fitted Scatterplot to check patterns in residuals before further analysis. -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resp-fit
#| fig-cap: "Response vs. Fitted plot to assess model fit."

y_hat <- fitted(model_1)
plot(x = y_hat, y = data$average_playtime, main = "Response vs. Fitted", xlab = "Fitted", ylab = "Average Playtime")
abline(a=0, b=1, lty=2)
```
To check the conditional mean response assumption, we analyze the Response vs. Fitted scatterplot @fig-resp-fit. The data points are tightly clustered in the bottom-left corner, which satisfies this assumption. The fitted line generally follows the line $y=hat{y}$, but there are notable outliers along the Y-axis where the observed `average_playtime` deviates significantly from the predicted values. These extreme outliers suggest that the response variable may be highly skewed or contain problematic data points.

To check the conditional mean predictor assumption, we examine the pairwise scatterplots of predictors @fig-pairwise. A lack of curves in all pairs of plots indicates that this assumption is satisfied. However, the Residuals vs. Fitted scatterplot @fig-resfit shows a fanning pattern resembling a left-pointing arrow, which suggests a potential violation of the constant variance assumption. Additionally, the Residuals vs. Qualitative Predictor scatterplots @fig-respre indicate possible violations of linearity and uncorrelated errors due to clusters of points and curves in the `positive_ratings` and `negative_ratings` predictors.

<!-- ## Create Pairwise Scatterplots for Predictors to look for multicollinearity -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pairwise
#| fig-cap: "Pairwise plot between numeric predictors"

ggpairs(data[, c(5, 6, 7, 11)]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8))

```

<!-- ## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resfit
#| fig-cap: "Residual vs fitted plot for model lm(average_playtime ~ price + positive_ratings + 
#|  negative_ratings + achievements + owners, data=data)"

y_value <- resid(model_1)
x_value <- fitted(model_1)

plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
abline(h = 0, col = "red", lty = 2)
```

Boxplots were used to assess linearity violations in the categorical predictor `owners`, and no issues were detected  @sec-model-detail. However, the Normal QQ Plot @fig-qq reveals significant deviations from the diagonal at the extremes, indicating a violation of the normality assumption.

The presence of these assumption violations aligns with the nature of the data, as confirmed by the outlier analysis in appendix @sec-outlier, which identifies several influential points even in the final model. For instance, the uncorrelated errors assumption may be violated because some predictors are inherently related (e.g., price and developer). These findings are consistent with those in the literature, where similar violations and errors have been reported. Addressing these violations through transformations is critical to ensure the stability of the model.

<!-- ## Create Normal QQ Plot to check Normality assumption -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-qq
#| fig-cap: "Residual vs fitted plot for model"

qqnorm(y_value)
qqline(y_value)
```
<!-- ## Box-Cox Transformation -->
```{r}
#| echo: false
#| warning: false
#| message: false

data$price <- data$price + 1
data$positive_ratings <- data$positive_ratings + 1
data$negative_ratings <- data$negative_ratings + 1
data$achievements <- data$achievements + 1

# Define the Box-Cox transformation function
boxcox_transform <- function(x) {
  bc <- BoxCoxTrans(x)
  predict(bc, x)
}
# Apply the transformation to each predictor and the response
bc_average_playtime <- boxcox_transform(data$average_playtime)
bc_price <- boxcox_transform(data$price)
bc_positive_ratings <- boxcox_transform(data$positive_ratings)
bc_negative_ratings <- boxcox_transform(data$negative_ratings)
bc_achievements <- boxcox_transform(data$achievements)

```
<!-- ## Fitting transformed model -->
```{r}
#| echo: false
#| warning: false
#| message: false

# Load bc transformed model
model_transformed <- readRDS(file = here::here("models/model_bc_trans.rds"))
```

To mitigate these violations, Box-Cox transformations were applied to all numeric predictors to improve normality and address linearity issues. Different transformations were applied to the response and predictors. For instance, the response variable `average_playtime` was log-transformed after calculating a Box-Cox $\lambda$ 0.1, as an example, an original average_playtime value of 17,612 was transformed to 9.78. Note that a constant of 1 was added to all predictors before applying the Box-Cox transformation, as it requires strictly positive values.

After transforming the model, the Variance Inflation Factor (VIF) was calculated to check for multicollinearity. Notably, `bc_positive_ratings` (log-transformed `positive_ratings`) and `bc_negative_ratings` (log-transformed `negative_ratings`) exhibited moderately high VIF values of approximately 5 and 4.5 respectively, indicating multicollinearity. This is reasonable, as a higher number of positive reviews typically corresponds to greater attention and a larger player base of a game, which can also result in more negative reviews. The pairwise scatterplots @fig-pairwise confirm this, showing a positive linear correlation between these two predictors.

<!-- ## Checking for Multicollinearity -->
```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

vif(model_transformed)
```
To address this issue, `bc_negative_ratings` was removed from the model, leaving the final predictors as `bc_price` (log-transformed of `price`), `bc_positive_ratings`(log-transformed of `positive_ratings`),  `bc_achievements`(log-transformed of `achievements`) and `owners`. The refined model was re-evaluated with summary statistics and residual plots. The summary of coefficients as shown in @tbl-summary will be discussed in Section @sec-conclusion. The residual diagnostics @fig-reduced show significant improvement, with a more random distribution around zero in the Residuals vs. Fitted plot and a QQ plot with better alignment to the reference line.

<!-- ## Generatingn Residual Plots -->
```{r, fig.width=8, fig.height=6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reduced
#| fig-cap: "Diagnostic plots for lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)"

reduced <- readRDS(file = here::here("models/model_bc_reduced.rds"))
par(mfrow = c(2, 2))
plot(reduced)
# y_value <- resid(reduced)
# x_value <- fitted(reduced)

# plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
```
To further simplify the model, automated selection procedures using AIC in both forward and backward directions were conducted. The procedure resulted in the same final model as described above, confirming its robustness.

<!-- ## AIC Model Selection -->
```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

step_model <- stepAIC(reduced, direction = "both")
summary(step_model)
```
<!-- ## ANOVA Comparison between full model and reduced model -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

anova(step_model, reduced)
# Conclude that step_model (without bc_achievements is better suited and no significant improvement to model from bc_achievements)
summary(step_model)$adj.r.squared
summary(reduced)$adj.r.squared
AIC(step_model)
AIC(reduced)
```

# Conclusion and Limitations {#sec-conclusion}

In conclusion, the model demonstrates that the average playtime of a game is positively influenced by several factors, including the game's price, the number of achievements, the number of positive reviews, and the estimated number of owners. Notably, the number of owners has the most substantial effect, as shown in the model summary @tbl-summary. For instance, games with an estimated ownership level of 100,000–200,000 have a coefficient of approximately 0.81; as ownership grows by a factor of 10, the effect increases to ~1.70, and for another 10-fold increase, the effect grows to ~3.04. This indicates that games with a larger player base tend to have higher average playtime, likely due to increased community engagement and multiplayer opportunities.

The positive relationship between price and average playtime is also significant. This is consistent with the notion that higher-priced games often provide more content, greater replayability, and better quality, which encourages players to spend more time playing. Additionally, players who invest more money in a game may feel a stronger obligation to "get their money's worth," resulting in longer playtime. Similarly, a greater number of positive reviews likely attracts more players and increases playtime, as positive feedback signals quality and encourages engagement.

Overall, the model aligns well with practical expectations. It suggests that developers and publishers aiming to increase average playtime should focus on improving game quality, fostering positive user reviews, and building a strong player base.

The limitations in this study is ...

# Ethics discussion
<!-- The Steam Games dataset [@davis2023steam] used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform. The metadata is comprehensively filled out, including details such as game titles, release dates, genres, and user ratings, which enhances the dataset's usability and reliability. The source of the data is clearly described, originating from the Steam store pages and Steam API, ensuring transparency and traceability. Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, and has been vetted and used in existing literature (discussed in the introduction). This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility through the framework outlined in the ethics module. There are no ethical concerns to report. -->

\newpage

\appendix

# Appendix {-}

# Model details {#sec-model-detail}
## Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-respre
#| fig-cap: "Residuals vs each Quantitative Predictor"

par(mfrow=c(1,4))

plot(x = data$price, y = y_value, main = "Residuals vs. Price", xlab = "Price", ylab = "Residuals")

plot(x = data$positive_ratings, y = y_value, main = "Residuals vs. Positive Ratings", xlab = "Positive Ratings", ylab = "Residuals")

plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")

plot(x = data$achievements, y = y_value, main = "Residuals vs. Achievements", xlab = "Achievements", ylab = "Residuals")
```

## BoxPlot to check assumptions for Categorical Predictors
See figure @fig-boc-cat
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-boc-cat
#| fig-cap: "Boxplot residual vs. Owner levels suggest no linear assumption violation regarding the categorical predictors"

# par(mfrow=c(1,2))

# boxplot(y_value ~ data$developer , main="Residual vs Developer",
# xlab="Developer", ylab="Residuals")

boxplot(y_value ~ data$owners , main="Residual vs Owners",
xlab="Owners", ylab="Residuals")
```
## Final model results
<!-- ## Fitting Reduced Model -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary

# Third model with reduce variable lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)
kable(tidy(reduced), caption = "Summary of the Reduced Model")
```

## Problematic points detection {#sec-outlier}

Checking for problematic points-leverage cutoff: ~0.006, cook's distance cutoff: ~0.961, dffits cutoff: ~0.105, dfbeta cutoff: ~0.025. Number of influential points by each case-leverage points: 216, regression outliers: 7, Cook's distance: 0, DFFITS: 211, by Beta from 0 to 7: 431, 476, 377, 394, 464, 321, 49, 0.
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

n <- nrow(data)
p <- length(coef(reduced)-1)
# leverage cutoff
hcut <- 2 * ((p + 1) / n)

# cook's distance cutoff
cookcut <- qf(0.5, df1 = p + 1, df2 = n - p - 1)

# dffits cutoff
fitcut <- 2 * sqrt((p + 1) / n)

# dfbeta cutoff
betacut <- 2 / sqrt(n)
# leverage statistic
h_ii <- hatvalues(reduced)

# standardized residuals
r_i <- rstandard(reduced)

# cook's distance 
D_i <- cooks.distance(reduced)

# dffits
dffits_i <- dffits(reduced)

# dfbetas
dfbetas_i <- dfbetas(reduced)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Count the number of leverage points
leverage_points <- which(h_ii > hcut)

# Count the number of regression outliers
regression_outliers <- which(r_i > 4 | r_i < -4)

# Count the number of influential observations by Cook's distance
influential_cooks <- which(D_i > cookcut)

# Count the number of influential observations by DFFITS
influential_dffits <- which(abs(dffits_i) > fitcut)

# Count the number of influential observations by DFBETAS
for (i in 1:8) {
  beta_influential <- which(abs(dfbetas_i[, i]) > betacut)
  cat(paste0("  Beta ", i - 1, ": ", length(beta_influential), "\n"))
}
```

\newpage

# References

### 2024-12-01 scripts/01-data_cleaning.R

@@ -13,7 +13,7 @@ library(readr)
 
 #### Clean data ####
 # Read the raw data
-raw_data <- read_csv("data/raw_data/raw_data.csv")
+raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"))
 
 # Clean and transform the data
 cleaned_data <- raw_data %>%

### 2024-12-01 scripts/02-model_data.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Save fitted model with analysis data 
+# Author: Ziqi Zhu
+# Date: 30 November 2024
+# Contact: ziqi.zhu@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: None
+
+
+#### Workspace setup ####
+library(tidyverse)
+#### preliminary data ####
+
+analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+
+# first model with all predictors
+model_1 <- lm(average_playtime ~ price + positive_ratings + negative_ratings + achievements + owners, data=analysis_data)
+
+# Second model with boc-cox transform on predictors
+analysis_data$price <- analysis_data$price + 1
+analysis_data$positive_ratings <- analysis_data$positive_ratings + 1
+analysis_data$negative_ratings <- analysis_data$negative_ratings + 1
+analysis_data$achievements <- analysis_data$achievements + 1
+
+# Define the Box-Cox transformation function
+boxcox_transform <- function(x) {
+  bc <- BoxCoxTrans(x)
+  predict(bc, x)
+}
+# Apply the transformation to each predictor and the response
+bc_average_playtime <- boxcox_transform(analysis_data$average_playtime)
+bc_price <- boxcox_transform(analysis_data$price)
+bc_positive_ratings <- boxcox_transform(analysis_data$positive_ratings)
+bc_negative_ratings <- boxcox_transform(analysis_data$negative_ratings)
+bc_achievements <- boxcox_transform(analysis_data$achievements)
+
+model_bc_trans <- lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements+ bc_negative_ratings + owners, data=analysis_data)
+
+# Third model with reduced predictor
+model_bc_reduced <- lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=analysis_data)
+# Save models as rds
+saveRDS(model_1, file = "models/model_1.rds")
+saveRDS(model_bc_trans, file = "models/model_bc_trans.rds")
+saveRDS(model_bc_reduced, file = "models/model_bc_reduced.rds")

### 2024-12-01 scripts/02-preliminary_data.R

@@ -1,11 +1,0 @@
-#### Preamble ####
-# Purpose: 
-# Author: 
-# Date: 7 October 2024
-# Contact: ziqi.zhu@mail.utoronto.ca 
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-
-#### preliminary data ####

### 2024-12-01 paper/paper.Rmd
fix a typo

@@
 The data points are tightly clustered in the bottom-left corner, which satisfies this assumption.
-The fitted line generally follows the line $y=hat{y}$, but there are notable outliers along the Y-axis where the observed `average_playtime` deviates significantly from the predicted values.
+The fitted line generally follows the line $y=\hat{y}$, but there are notable outliers along the Y-axis where the observed `average_playtime` deviates significantly from the predicted values.
 These extreme outliers suggest that the response variable may be highly skewed or contain problematic data points.

### 2024-12-04 paper/paper.Rmd
Rendered paper into pdf.

@@
     colorlinks: true
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 # Contributions
-We found the dataset together, and seperate the works as follow:
-Ziqi: Introduction and data description
-Raghav: Ethics discussion and some plot in preliminary result
-Christophe: Most of Preliminary results section
+
+We found the dataset together, and seperate the works as follow: Ziqi:
+Introduction and data description Raghav: Ethics discussion and some
+plot in preliminary result Christophe: Most of Preliminary results
+section
 
 # Introduction
-The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam.
-Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers.
-Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts.
-So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle.
-Specifically, we will investigate how variables such as price, user ratings, and other game characteristics contribute to player engagement, as measured by playtime.
-From some previous papers, we learned that both positive and negative reviews have a significant correlation with playtime, but the effect is more pronounced for positive reviews [@paper1].
-Another paper found out that game pricing significantly affects player engagement, with lower-priced games generally seeing higher playtime [@paper2].
-The third paper did an empirical study that also examined the influence of game reviews on playtime, focusing on how the number and sentiment of reviews can affect player engagement.
-It concludes that games with more positive reviews see increased playtime, especially if those reviews highlight the game’s quality [@paper3].
-We found this problem suitable for linear regression because it allows us to model the relationship between average playtime (the dependent variable) and various predictors (price, ratings, developer, etc.).
-This approach helps quantify how each factor influences playtime, enabling a clear understanding of the impact of different game attributes.
-By fitting a linear regression model, we can determine which factors have statistically significant effects on playtime, providing actionable insights for game developers and marketers.
+
+The gaming industry has experienced exponential growth in recent years,
+with millions of players engaging in various games on platforms like
+Steam. Understanding what influences average playtime in games is
+crucial for game developers, publishers, and marketers. Average playtime
+is a key indicator of player engagement and game success, and
+investigating factors that impact it can provide valuable insights into
+game design, pricing strategies, and marketing efforts. So our research
+aims to explore the factors that influence the average playtime of games
+available on Steam using a dataset we get from Kaggle. Specifically, we
+will investigate how variables such as price, user ratings, and other
+game characteristics contribute to player engagement, as measured by
+playtime. From some previous papers, we learned that both positive and
+negative reviews have a significant correlation with playtime, but the
+effect is more pronounced for positive reviews [@paper1]. Another paper
+found out that game pricing significantly affects player engagement,
+with lower-priced games generally seeing higher playtime [@paper2]. The
+third paper did an empirical study that also examined the influence of
+game reviews on playtime, focusing on how the number and sentiment of
+reviews can affect player engagement. It concludes that games with more
+positive reviews see increased playtime, especially if those reviews
+highlight the game’s quality [@paper3]. We found this problem suitable
+for linear regression because it allows us to model the relationship
+between average playtime (the dependent variable) and various predictors
+(price, ratings, developer, etc.). This approach helps quantify how each
+factor influences playtime, enabling a clear understanding of the impact
+of different game attributes. By fitting a linear regression model, we
+can determine which factors have statistically significant effects on
+playtime, providing actionable insights for game developers and
+marketers.
 
@@
 
-To develop and refine the final model for predicting average playtime, a structured approach was employed that combined diagnostic assessments, transformations, and model selection techniques.
-The analysis began with data preparation and selection of relevant predictors based on domain knowledge and preliminary exploration.
-A multiple linear regression model was initially fitted using predictors including price, positive ratings, negative ratings, number of achievements as numeric variables, and ownership levels as categorical variable.
+To develop and refine the final model for predicting average playtime, a
+structured approach was employed that combined diagnostic assessments,
+transformations, and model selection techniques. The analysis began with
+data preparation and selection of relevant predictors based on domain
+knowledge and preliminary exploration. A multiple linear regression
+model was initially fitted using predictors including price, positive
+ratings, negative ratings, number of achievements as numeric variables,
+and ownership levels as categorical variable.
 
-Preliminary model diagnostics, since we are using MLR model, we first need to check conditional mean condition with both response and predictors, including response vs.
-fitted plots and pairwise plots between predictors.
-Then residuals vs.
-fitted plots were used to assess the assumptions of linearity, constant-variance, and uncorrelated errors.
-While Normal Quantile-Quantile (QQ) plot is used to check for normality.
-To address violations of these assumptions, Box-Cox transformations were applied to the response variable and select predictors to make the most suited transform function based on their $\lambda$.
-Then pairwise scatterplots and the Variance Inflation Factor (VIF) were used to check for multicollinearity, and highly correlated variables were removed or adjusted to improve the model's stability.
+Preliminary model diagnostics, since we are using MLR model, we first
+need to check conditional mean condition with both response and
+predictors, including response vs. fitted plots and pairwise plots
+between predictors. Then residuals vs. fitted plots were used to assess
+the assumptions of linearity, constant-variance, and uncorrelated
+errors. While Normal Quantile-Quantile (QQ) plot is used to check for
+normality. To address violations of these assumptions, Box-Cox
+transformations were applied to the response variable and select
+predictors to make the most suited transform function based on their
+$\lambda$. Then pairwise scatterplots and the Variance Inflation Factor
+(VIF) were used to check for multicollinearity, and highly correlated
+variables were removed or adjusted to improve the model's stability.
 
-Refinement of the model was carried out through stepwise selection using AIC (Akaike Information Criterion) to identify the best combination of predictors while balancing model complexity and fit.
-ANOVA tests were used to compare nested models and evaluate whether reductions in predictors led to significant changes in explanatory power.
-The final model was validated using diagnostic plots to confirm adherence to linear regression assumptions, ensuring that it was suitable for interpretation and prediction.
+Refinement of the model was carried out through stepwise selection using
+AIC (Akaike Information Criterion) to identify the best combination of
+predictors while balancing model complexity and fit. ANOVA tests were
+used to compare nested models and evaluate whether reductions in
+predictors led to significant changes in explanatory power. The final
+model was validated using diagnostic plots to confirm adherence to
+linear regression assumptions, ensuring that it was suitable for
+interpretation and prediction.
 
 The methodology was implemented in programming language **R** [@citeR].
-Packages **readr** [@readr], **httr** [@httr], **jsonlite** [@jsonlite] were used to download the data, **dplyr** [@dplyr] and **lubridate** [@lubridate] were used to clean the data.
-The diagnostic process is done with help of **ggplot2** [@ggplot2] for residual graphs, **GGally** [@GGally] and **car** [@car] for multicollinearity checks, **MASS** [@MASS] for stepwise selection, **caret** [@caret] is used to perform box-cox transform on variables.
-This systematic approach ensured that each decision was guided by statistical evidence and diagnostic tools, allowing for the development of a robust final model.
-A simplified version of the methodology flowchart is presented in @fig-flowchart.
+Packages **readr** [@readr], **httr** [@httr], **jsonlite** [@jsonlite]
+were used to download the data, **dplyr** [@dplyr] and **lubridate**
+[@lubridate] were used to clean the data. The diagnostic process is done
+with help of **ggplot2** [@ggplot2] for residual graphs, **GGally**
+[@GGally] and **car** [@car] for multicollinearity checks, **MASS**
+[@MASS] for stepwise selection, **caret** [@caret] is used to perform
+box-cox transform on variables. This systematic approach ensured that
+each decision was guided by statistical evidence and diagnostic tools,
+allowing for the development of a robust final model. A simplified
+version of the methodology flowchart is presented in @fig-flowchart.
 
@@
 <!-- ## Fit multiple linear regression model -->
+
 ```{r}
@@
 ```
+
 # Results
 
-At the starting point, our baseline model uses `average_playtime` as the response variable, with `price`, `positive_ratings`, `negative_ratings`, `achievements`, and `owners` as predictors.
-The four key linear regression assumptions considered are linearity, uncorrelated errors, constant error variance, and normality.
-Additionally, since this is a multiple linear regression model, we also examine the conditional mean response and predictor assumptions.
+At the starting point, our baseline model uses `average_playtime` as the
+response variable, with `price`, `positive_ratings`, `negative_ratings`,
+`achievements`, and `owners` as predictors. The four key linear
+regression assumptions considered are linearity, uncorrelated errors,
+constant error variance, and normality. Additionally, since this is a
+multiple linear regression model, we also examine the conditional mean
+response and predictor assumptions.
 <!-- ## Create Response vs Fitted Scatterplot to check patterns in residuals before further analysis. -->
+
 ```{r}
@@
 ```
-To check the conditional mean response assumption, we analyze the Response vs.
-Fitted scatterplot @fig-resp-fit.
-The data points are tightly clustered in the bottom-left corner, which satisfies this assumption.
-The fitted line generally follows the line $y=\hat{y}$, but there are notable outliers along the Y-axis where the observed `average_playtime` deviates significantly from the predicted values.
-These extreme outliers suggest that the response variable may be highly skewed or contain problematic data points.
 
-To check the conditional mean predictor assumption, we examine the pairwise scatterplots of predictors @fig-pairwise.
-A lack of curves in all pairs of plots indicates that this assumption is satisfied.
-However, the Residuals vs.
-Fitted scatterplot @fig-resfit shows a fanning pattern resembling a left-pointing arrow, which suggests a potential violation of the constant variance assumption.
-Additionally, the Residuals vs.
-Qualitative Predictor scatterplots @fig-respre indicate possible violations of linearity and uncorrelated errors due to clusters of points and curves in the `positive_ratings` and `negative_ratings` predictors.
+To check the conditional mean response assumption, we analyze the
+Response vs. Fitted scatterplot @fig-resp-fit. The data points are
+tightly clustered in the bottom-left corner, which satisfies this
+assumption. The fitted line generally follows the line $y=\hat{y}$, but
+there are notable outliers along the Y-axis where the observed
+`average_playtime` deviates significantly from the predicted values.
+These extreme outliers suggest that the response variable may be highly
+skewed or contain problematic data points.
+
+To check the conditional mean predictor assumption, we examine the
+pairwise scatterplots of predictors @fig-pairwise. A lack of curves in
+all pairs of plots indicates that this assumption is satisfied. However,
+the Residuals vs. Fitted scatterplot @fig-resfit shows a fanning pattern
+resembling a left-pointing arrow, which suggests a potential violation
+of the constant variance assumption. Additionally, the Residuals vs.
+Qualitative Predictor scatterplots @fig-respre indicate possible
+violations of linearity and uncorrelated errors due to clusters of
+points and curves in the `positive_ratings` and `negative_ratings`
+predictors.
 
 <!-- ## Create Pairwise Scatterplots for Predictors to look for multicollinearity -->
+
 ```{r}
@@
 <!-- ## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance -->
+
 ```{r}
@@
 
-Boxplots were used to assess linearity violations in the categorical predictor `owners`, and no issues were detected  @sec-model-detail.
-However, the Normal QQ Plot @fig-qq reveals significant deviations from the diagonal at the extremes, indicating a violation of the normality assumption.
+Boxplots were used to assess linearity violations in the categorical
+predictor `owners`, and no issues were detected @sec-model-detail.
+However, the Normal QQ Plot @fig-qq reveals significant deviations from
+the diagonal at the extremes, indicating a violation of the normality
+assumption.
 
-The presence of these assumption violations aligns with the nature of the data, as confirmed by the outlier analysis in appendix @sec-outlier, which identifies several influential points even in the final model.
-For instance, the uncorrelated errors assumption may be violated because some predictors are inherently related (e.g., price and developer).
-These findings are consistent with those in the literature, where similar violations and errors have been reported.
-Addressing these violations through transformations is critical to ensure the stability of the model.
+The presence of these assumption violations aligns with the nature of
+the data, as confirmed by the outlier analysis in appendix @sec-outlier,
+which identifies several influential points even in the final model. For
+instance, the uncorrelated errors assumption may be violated because
+some predictors are inherently related (e.g., price and developer).
+These findings are consistent with those in the literature, where
+similar violations and errors have been reported. Addressing these
+violations through transformations is critical to ensure the stability
+of the model.
 
 <!-- ## Create Normal QQ Plot to check Normality assumption -->
+
 ```{r}
@@
 ```
+
 <!-- ## Box-Cox Transformation -->
+
 ```{r}
@@
 ```
+
 <!-- ## Fitting transformed model -->
+
 ```{r}
@@
 
-To mitigate these violations, Box-Cox transformations were applied to all numeric predictors to improve normality and address linearity issues.
-Different transformations were applied to the response and predictors.
-For instance, the response variable `average_playtime` was log-transformed after calculating a Box-Cox $\lambda$ 0.1, as an example, an original average_playtime value of 17,612 was transformed to 9.78.
-Note that a constant of 1 was added to all predictors before applying the Box-Cox transformation, as it requires strictly positive values.
+To mitigate these violations, Box-Cox transformations were applied to
+all numeric predictors to improve normality and address linearity
+issues. Different transformations were applied to the response and
+predictors. For instance, the response variable `average_playtime` was
+log-transformed after calculating a Box-Cox $\lambda$ 0.1, as an
+example, an original average_playtime value of 17,612 was transformed to
+9.78. Note that a constant of 1 was added to all predictors before
+applying the Box-Cox transformation, as it requires strictly positive
+values.
 
-After transforming the model, the Variance Inflation Factor (VIF) was calculated to check for multicollinearity.
-Notably, `bc_positive_ratings` (log-transformed `positive_ratings`) and `bc_negative_ratings` (log-transformed `negative_ratings`) exhibited moderately high VIF values of approximately 5 and 4.5 respectively, indicating multicollinearity.
-This is reasonable, as a higher number of positive reviews typically corresponds to greater attention and a larger player base of a game, which can also result in more negative reviews.
-The pairwise scatterplots @fig-pairwise confirm this, showing a positive linear correlation between these two predictors.
+After transforming the model, the Variance Inflation Factor (VIF) was
+calculated to check for multicollinearity. Notably,
+`bc_positive_ratings` (log-transformed `positive_ratings`) and
+`bc_negative_ratings` (log-transformed `negative_ratings`) exhibited
+moderately high VIF values of approximately 5 and 4.5 respectively,
+indicating multicollinearity. This is reasonable, as a higher number of
+positive reviews typically corresponds to greater attention and a larger
+player base of a game, which can also result in more negative reviews.
+The pairwise scatterplots @fig-pairwise confirm this, showing a positive
+linear correlation between these two predictors.
 
 <!-- ## Checking for Multicollinearity -->
+
 ```{r}
@@
 ```
-To address this issue, `bc_negative_ratings` was removed from the model, leaving the final predictors as `bc_price` (log-transformed of `price`), `bc_positive_ratings`(log-transformed of `positive_ratings`),  `bc_achievements`(log-transformed of `achievements`) and `owners`.
-The refined model was re-evaluated with summary statistics and residual plots.
-The summary of coefficients as shown in @tbl-summary will be discussed in Section @sec-conclusion.
-The residual diagnostics @fig-reduced show significant improvement, with a more random distribution around zero in the Residuals vs.
-Fitted plot and a QQ plot with better alignment to the reference line.
+
+To address this issue, `bc_negative_ratings` was removed from the model,
+leaving the final predictors as `bc_price` (log-transformed of `price`),
+`bc_positive_ratings`(log-transformed of `positive_ratings`),
+`bc_achievements`(log-transformed of `achievements`) and `owners`. The
+refined model was re-evaluated with summary statistics and residual
+plots. The summary of coefficients as shown in @tbl-summary will be
+discussed in Section @sec-conclusion. The residual diagnostics
+@fig-reduced show significant improvement, with a more random
+distribution around zero in the Residuals vs. Fitted plot and a QQ plot
+with better alignment to the reference line.
 
 <!-- ## Generatingn Residual Plots -->
+
 ```{r, fig.width=8, fig.height=6}
@@
 ```
-To further simplify the model, automated selection procedures using AIC in both forward and backward directions were conducted.
-The procedure resulted in the same final model as described above, confirming its robustness.
+
+To further simplify the model, automated selection procedures using AIC
+in both forward and backward directions were conducted. The procedure
+resulted in the same final model as described above, confirming its
+robustness.
 
 <!-- ## AIC Model Selection -->
+
 ```{r}
@@
 ```
+
 <!-- ## ANOVA Comparison between full model and reduced model -->
+
 ```{r}
@@
 
-In conclusion, the model demonstrates that the average playtime of a game is positively influenced by several factors, including the game's price, the number of achievements, the number of positive reviews, and the estimated number of owners.
-Notably, the number of owners has the most substantial effect, as shown in the model summary @tbl-summary.
-For instance, games with an estimated ownership level of 100,000–200,000 have a coefficient of approximately 0.81; as ownership grows by a factor of 10, the effect increases to ~1.70, and for another 10-fold increase, the effect grows to ~3.04.
-This indicates that games with a larger player base tend to have higher average playtime, likely due to increased community engagement and multiplayer opportunities.
+In conclusion, the model demonstrates that the average playtime of a
+game is positively influenced by several factors, including the game's
+price, the number of achievements, the number of positive reviews, and
+the estimated number of owners. Notably, the number of owners has the
+most substantial effect, as shown in the model summary @tbl-summary. For
+instance, games with an estimated ownership level of 100,000–200,000
+have a coefficient of approximately 0.81; as ownership grows by a factor
+of 10, the effect increases to \~1.70, and for another 10-fold increase,
+the effect grows to \~3.04. This indicates that games with a larger
+player base tend to have higher average playtime, likely due to
+increased community engagement and multiplayer opportunities.
 
-The positive relationship between price and average playtime is also significant.
-This is consistent with the notion that higher-priced games often provide more content, greater replayability, and better quality, which encourages players to spend more time playing.
-Additionally, players who invest more money in a game may feel a stronger obligation to "get their money's worth," resulting in longer playtime.
-Similarly, a greater number of positive reviews likely attracts more players and increases playtime, as positive feedback signals quality and encourages engagement.
+The positive relationship between price and average playtime is also
+significant. This is consistent with the notion that higher-priced games
+often provide more content, greater replayability, and better quality,
+which encourages players to spend more time playing. Additionally,
+players who invest more money in a game may feel a stronger obligation
+to "get their money's worth," resulting in longer playtime. Similarly, a
+greater number of positive reviews likely attracts more players and
+increases playtime, as positive feedback signals quality and encourages
+engagement.
 
-Overall, the model aligns well with practical expectations.
-It suggests that developers and publishers aiming to increase average playtime should focus on improving game quality, fostering positive user reviews, and building a strong player base.
+Overall, the model aligns well with practical expectations. It suggests
+that developers and publishers aiming to increase average playtime
+should focus on improving game quality, fostering positive user reviews,
+and building a strong player base.
 
-The limitations in this study is ...
+This study faces limitations primarily due to assumption violations and
+data issues. The Residuals vs. Fitted plot revealed a fanning pattern,
+indicating heteroscedasticity, while the Normal QQ Plot showed
+significant non-normality. Outliers in the response variable
+`average_playtime` suggest skewed data, despite transformations like
+Box-Cox improving stability. Multicollinearity between
+`bc_positive_ratings` and `bc_negative_ratings` required removing the
+latter, though some interdependence remains.
+
+The dataset, limited to Steam games, introduces platform-specific
+biases, restricting generalizability. Observational data prevents causal
+inference, and omitted variables, such as marketing strategies, likely
+impact playtime. Future studies should refine transformations, address
+multicollinearity more rigorously, and expand variables and platforms to
+enhance the robustness and applicability of the findings.
 
 # Ethics discussion
-<!-- The Steam Games dataset [@davis2023steam] used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform.
-The metadata is comprehensively filled out, including details such as game titles, release dates, genres, and user ratings, which enhances the dataset's usability and reliability.
-The source of the data is clearly described, originating from the Steam store pages and Steam API, ensuring transparency and traceability.
-Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, and has been vetted and used in existing literature (discussed in the introduction).
-This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility through the framework outlined in the ethics module.
-There are no ethical concerns to report.
--->
+
+The Steam Games dataset [@davis2023steam] used in this research contains
+metadata on over 97,000 games published on the Steam platform. The
+dataset is sourced transparently from Steam store pages and the Steam
+API, hosted on Kaggle, and widely used in existing literature, ensuring
+its credibility.
+
+In our analysis, we chose automated selection methods due to the
+dataset's size and complexity. Manual selection methods were not used as
+they would have been impractical and prone to human error, reducing
+consistency and reproducibility. While both methods are ethically
+equivalent when executed carefully, automated methods better align with
+the ethical principles outlined in the second ethics module by
+minimizing foreseeable risks and ensuring reliability.
+
+Avoiding blameworthy practices requires addressing foreseeability,
+avoidability, and potential errors. Automated methods allowed us to
+mitigate these risks efficiently, ensuring accountability in our
+approach. Manual methods, while suitable for smaller datasets, would
+have introduced avoidable errors and diminished reliability, which could
+be considered ethically negligent.
+
+By prioritizing virtues such as precision and accountability, we ensured
+an ethically sound methodology. Our choice was guided by practical
+considerations without compromising ethical standards, fostering robust
+and reproducible results.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
 # Model details {#sec-model-detail}
+
 ## Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
+
 ```{r}
@@
 ## BoxPlot to check assumptions for Categorical Predictors
+
 See figure @fig-boc-cat
+
 ```{r}
@@
 ```
+
 ## Final model results
+
 <!-- ## Fitting Reduced Model -->
+
 ```{r}
@@
 
-Checking for problematic points-leverage cutoff: ~0.006, cook's distance cutoff: ~0.961, dffits cutoff: ~0.105, dfbeta cutoff: ~0.025.
-Number of influential points by each case-leverage points: 216, regression outliers: 7, Cook's distance: 0, DFFITS: 211, by Beta from 0 to 7: 431, 476, 377, 394, 464, 321, 49, 0.
+Checking for problematic points-leverage cutoff: \~0.006, cook's
+distance cutoff: \~0.961, dffits cutoff: \~0.105, dfbeta cutoff:
+\~0.025. Number of influential points by each case-leverage points: 216,
+regression outliers: 7, Cook's distance: 0, DFFITS: 211, by Beta from 0
+to 7: 431, 476, 377, 394, 464, 321, 49, 0.
+
 ```{r}

### 2024-12-04 paper/editing.Rmd
Edited the introduction to increase elegance and clarity, and added editing process document.

---
title: "Editing Process Reflection"
author: 
  - name: "Raghav Sinha"
  - name: "Ziqi Zhu"
  - name: "Chris McWatt"
date: "`r Sys.Date()`"
date-format: long
abstract: ""
number-sections: false
format:
  pdf:
    toc: false
    number-sections: false
    colorlinks: true
editor: 
  markdown: 
    wrap: 72
---

# Editing Process Reflection

## Selected Editing Technique
- **Technique:** Word Choice and Ordering (concrete nouns, strong verbs)
- **Reason:** By carefully choosing nouns and verbs, the sentence structure becomes tighter, reducing wordiness and redundancy. This allows the writing to communicate the message more efficiently, maintaining clarity while also elevating the tone.
- **Focus Area:** Simplicity, Elegance

## Original Text
> The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam. Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers. Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts. So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle.

## Edited Text
> The gaming industry has experienced remarkable growth in recent years, with millions of players engaging on platforms like Steam. Understanding the factors that influence average playtime is crucial for developers, publishers, and marketers alike. As a key indicator of player engagement and game success, playtime can offer valuable insights into game design, pricing, and marketing strategies. Our research aims to explore the factors that affect the average playtime of games on Steam, using a dataset sourced from Kaggle. Specifically, we will examine how variables such as price, user ratings, and other game attributes contribute to player engagement.

## Reflections on the Editing Process
- **What Went Well:** Redundant phrases were removed, and word choice and ordering were refined, resulting in a more streamlined and efficient expression of the motivations of our study without losing essential content.
- **Challenges:** Editing for elegance and simplicity without oversimplifying or losing the technical precision needed for statistical analyses, especially when dealing with specific research terms and concepts, was a delicate task.

### 2024-12-04 paper/paper.Rmd
Edited the introduction to increase elegance and clarity, and added editing process document.

@@
 
-The gaming industry has experienced exponential growth in recent years,
-with millions of players engaging in various games on platforms like
-Steam. Understanding what influences average playtime in games is
-crucial for game developers, publishers, and marketers. Average playtime
-is a key indicator of player engagement and game success, and
-investigating factors that impact it can provide valuable insights into
-game design, pricing strategies, and marketing efforts. So our research
-aims to explore the factors that influence the average playtime of games
-available on Steam using a dataset we get from Kaggle. Specifically, we
-will investigate how variables such as price, user ratings, and other
-game characteristics contribute to player engagement, as measured by
-playtime. From some previous papers, we learned that both positive and
-negative reviews have a significant correlation with playtime, but the
-effect is more pronounced for positive reviews [@paper1]. Another paper
-found out that game pricing significantly affects player engagement,
-with lower-priced games generally seeing higher playtime [@paper2]. The
-third paper did an empirical study that also examined the influence of
-game reviews on playtime, focusing on how the number and sentiment of
-reviews can affect player engagement. It concludes that games with more
-positive reviews see increased playtime, especially if those reviews
-highlight the game’s quality [@paper3]. We found this problem suitable
-for linear regression because it allows us to model the relationship
-between average playtime (the dependent variable) and various predictors
-(price, ratings, developer, etc.). This approach helps quantify how each
-factor influences playtime, enabling a clear understanding of the impact
-of different game attributes. By fitting a linear regression model, we
-can determine which factors have statistically significant effects on
-playtime, providing actionable insights for game developers and
-marketers.
+The gaming industry has experienced remarkable growth in recent years, with millions of players engaging on platforms like Steam.
+Understanding the factors that influence average playtime is crucial for developers, publishers, and marketers alike.
+As a key indicator of player engagement and game success, playtime can offer valuable insights into game design, pricing, and marketing strategies.
+Our research aims to explore the factors that affect the average playtime of games on Steam, using a dataset sourced from Kaggle.
+Specifically, we will examine how variables such as price, user ratings, and other game attributes contribute to player engagement.
+Previous studies have shown that both positive and negative reviews correlate with playtime, though the effect of positive reviews is more pronounced [@paper1].
+Additionally, research suggests that pricing plays a significant role, with lower-priced games generally seeing higher playtime [@paper2].
+Another study highlighted the influence of review sentiment, concluding that games with more positive reviews tend to have increased playtime, especially when those reviews emphasize game quality [@paper3].
+We have chosen linear regression to model the relationship between playtime (the dependent variable) and various predictors, such as price and ratings.
+This approach will quantify how each factor impacts playtime, providing clear insights into the influence of game attributes.
+By fitting this model, we aim to identify the factors with statistically significant effects on playtime, offering actionable recommendations for developers and marketers.
 

### 2024-12-06 paper/paper.Rmd
update contributions and readme

@@
 # Contributions
+Ziqi: Collaborated on the report, providing detailed explanations of methods and results. Organized files and scripts for the project.
 
-We found the dataset together, and seperate the works as follow: Ziqi:
-Introduction and data description Raghav: Ethics discussion and some
-plot in preliminary result Christophe: Most of Preliminary results
-section
+Raghav: Collaborated on the report and conducted early-stage analysis following the flowchart, including coding for each step.
+Addressed ethical discussions and completed the editing requirements workshop sheet.
+
+Christophe: Converted the report into a poster deliverable.
+
+Peer Review: Each member reviewed the work of others to ensure quality and consistency.
 
@@
 As a key indicator of player engagement and game success, playtime can offer valuable insights into game design, pricing, and marketing strategies.
-Our research aims to explore the factors that affect the average playtime of games on Steam, using a dataset sourced from Kaggle.
+Our research aims to explore the factors that affect the average playtime of games on Steam, using a dataset sourced from Kaggle [@davis2023steam].
 Specifically, we will examine how variables such as price, user ratings, and other game attributes contribute to player engagement.
@@
 Another study highlighted the influence of review sentiment, concluding that games with more positive reviews tend to have increased playtime, especially when those reviews emphasize game quality [@paper3].
-We have chosen linear regression to model the relationship between playtime (the dependent variable) and various predictors, such as price and ratings.
+We have chosen linear regression to model the relationship between playtime (the dependent variable) and several predictors, including price, positive and negative ratings, the number of achievements set up in the game, and the estimated number of game owners.
 This approach will quantify how each factor impacts playtime, providing clear insights into the influence of game attributes.

### 2024-12-06 paper/editing.Rmd
Reduced font size and changed documentclass to article.

@@
   pdf:
+    documentclass: article
     toc: false
@@
     colorlinks: true
+    margin-top: 0.75in
+    margin-bottom: 0.75in
+    margin-left: 0.75in
+    margin-right: 0.75in
 editor: 
